---
title: "Intro to Causality"
subtitle: "Election Data Science"  
author: 
  - "Peter Licari"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF",
  text_font_google = google_font("Poppins")
)

xaringanExtra::use_xaringan_extra()
library(tidyverse)
library(kableExtra)
```

```{css, echo=FALSE}
pre code, pre, code {
  white-space: pre !important;
  height: 100px !important;}
```
--


*Passphrase today is: Cause and Effect*

---
# What do we mean by "causal"?
--

**There is no single, set theory of causality. Instead there is a constellation of concepts that we glomm together when we discuss the topic:**
--
1. Deterministic causality.
2. Probabilistic causality.
3. Procedural causality.
4. Necessary/Sufficient causality.\*
5. Leverage causality.
6. Reciprocal causality.

--
**_In general_** there are three requirements to a causal relationship across types: There is a relationship between cause and effect; cause must precede effect; and there must not be any spurious causes.

---

Probabilistic causality: $X$ can be said to cause $Y$ if changes in $X$ significantly alters *the probability* of $Y$ occurring (provided the main three requirements are met). 

---
# What do we mean by significant.


---
# 4 concepts we need to understand to get statistical significance:

1. Statements of significance are always *statements of comparison.*
2. Most frequently, the comparison is to a hypothesized **null** relationship.
  - *That is, there is no difference between groups and/or no relationship between $X$ and $Y$.*
3. Determinations of "significance" is arbitrary.
  - The underlying concept is probabilistic; the up/down decision is based on your tolerance for uncertainty.
4. Significance **alone** cannot tell you the *importance* of the relationship. 

---
# Let's think about comparisons between groups. 
.panelset[.panel[.panel-name[Comparison 1]

```{r, echo=FALSE}
tibble(group1 = rnorm(100, mean = 0, sd = 5), group2 = rnorm(100, mean = 25, sd = 5)) %>%
  mutate(blah = "blah") %>%
  pivot_longer(!blah, names_to = "groups") %>%
  ggplot()+geom_density(aes(x=value, fill = groups), alpha = .6) + 
  theme_minimal() + xlab("")+ 
  theme(axis.text.x = element_blank()) 
```
]
.panel[.panel-name[Comparison 2]
```{r, echo=FALSE}
tibble(group1 = rnorm(100, mean = 0, sd = 5), group2 = rnorm(100, mean = 15, sd = 5)) %>%
  mutate(blah = "blah") %>%
  pivot_longer(!blah, names_to = "groups") %>%
  ggplot()+geom_density(aes(x=value, fill = groups), alpha = .6) + 
  theme_minimal() + xlab("")+ 
  theme(axis.text.x = element_blank()) 
```
]
.panel[.panel-name[Comparison 3]
```{r, echo=FALSE}
tibble(group1 = rnorm(100, mean = 0, sd = 5), group2 = rnorm(100, mean = 1, sd = 5)) %>%
  mutate(blah = "blah") %>%
  pivot_longer(!blah, names_to = "groups") %>%
  ggplot()+geom_density(aes(x=value, fill = groups), alpha = .6) + 
  theme_minimal() + xlab("")+ 
  theme(axis.text.x = element_blank()) 
```
]

]

---
# What information do we need?
.panelset[.panel[.panel-name[Central Tendency]
```{r}
tibble(group1 = rnorm(100, mean = 0, sd = 5), group2 = rnorm(100, mean = 25, sd = 5)) %>%
  mutate(blah = "blah") %>%
  pivot_longer(!blah, names_to = "groups") %>%
  ggplot()+geom_density(aes(x=value, fill = groups), alpha = .6) + 
  geom_segment(aes(x = 25, xend = 25, y = 0, yend = .1), color = "#2D4396")+
  geom_segment(aes(x = 0, xend = 0, y = 0, yend = .1), color = "#932C2C")+
  theme_minimal() + xlab("")+ 
  theme(axis.text.x = element_blank()) 
```

]
.panel[.panel-name[Distribution]
```{r}
tibble(group1 = rnorm(1000, mean = 0, sd = 5), group2 = rnorm(1000, mean = 25, sd = 5)) %>%
  mutate(blah = "blah") %>%
  pivot_longer(!blah, names_to = "groups") %>%
  ggplot()+geom_density(aes(x=value, fill = groups), alpha = .6) + 
  geom_segment(aes(x = 20, xend = 30, y = .0375, yend = .0375), color = "#2D4396")+
  geom_segment(aes(x = -5, xend = 5, y = .0375, yend = .0375), color = "#932C2C")+
  theme_minimal() + xlab("")+ 
  theme(axis.text.x = element_blank()) 
```


]

.panel[.panel-name[Quality]
Quality consists of the number of people you've selected and (implicitly) how they were gathered.
--

This tells us how **reliably** we can take the difference observed between the two groups.

**In our case:_Random sample of 1,000 members of both group 1 and group 2._**
]
]
---
# How sure is "sure"?
With all of this data, you can calculate a specific parameter ($t$ statistic) that will help you calculate a $p$ value. 

The p value represents how probable it is that the difference you observed, given the quality of your samples, would pop up if the *null* hypothesis (e.g., that **there is _no_ difference between the two groups** were true.) 

---

```{r, echo=FALSE}
df <- tibble()
for(i in seq_along(1:50)){
x <- tibble(group1 = rnorm(30, mean = 0, sd = 1), group2 = rnorm(30, mean = 0, sd = 1)) %>%
  mutate(blah = "blah") %>%
  pivot_longer(!blah, names_to = "groups") %>%
    mutate(z = i)
df<- rbind(df,x)
  }
ggplot(data = df) + geom_density(aes(x=value, fill = groups), alpha = .6) + theme_bw()+
  theme(axis.text = element_blank())+facet_wrap(~z, ncol = 10)
```

---

- The $p$ value tells you the probability that you would see a difference this extreme if the null hypothesis were true. 

- Statistical significance is determined based off of how often, on average, you (or, more realistically, your field), is comfortable accepting a so-called false-positive. 
  - Frequently see 5% of the time ($p = 0.05$) or 10% of the time ($p = 0.10$)
  
---
## Linear Regression. 

In order to understand the logic of linear regression, you really don't need to understand *much* more than high school algebra. Specifically:

<center>$Y = mX + b$</center>

---
## Algebra refresher:

```{r, echo=FALSE}
tibble(x = c(-5:5), y = c(-3:7)) %>%
  ggplot() + geom_line(aes(x = x, y = y), color = "blue") +
  geom_segment(aes(x = -5, xend = 5, y = 0, yend=0)) + 
  geom_segment(aes(x = 0, xend = 0, y = -7.5, yend=7.5))+
  ylim(c(-7.5, 7.5)) + xlim(c(-5,5))
```


---

