---
title: "Data Visualization with ggplot"
subtitle: "Election Data Science"  
author: 
  - "Peter Licari"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF",
  text_font_google = google_font("Poppins")
)

xaringanExtra::use_xaringan_extra()
library(tidyverse)
library(kableExtra)
```


# What we'll be learning:
--

- The 3 main chart types (5 `geom_`s)
  1. Bar charts (`geom_bar` and `geom_hist`)
  2. Dot charts (`geom_dot`)
  3. Line charts (`geom_line` and `geom_smooth`)
--

- Different aesthetic options
  1. Color/Fill
  2. Size
  3. Transparency
  4. Positioning/Grouping
  
--

- Some helpful theme stuff
  
---
# Let's grab the data:

```{r}
library(tidyverse)
library(anesr)

# 2016 American National Election Survey
data(timeseries_2016)
anes16 <- timeseries_2016
rm(timeseries_2016)
```

---
# First, some semantics.

```{r, eval=FALSE}

ggplot(data, aes(x=x, y=y, color = z)) + geom_point()

ggplot() + geom_point(data, aes(x=x, y=y, color = z))

ggplot(data) + geom_point(aes(x=x, y=y, color = z))
```

<center> **These will all make the same images!**


---

# Bar Charts (`geom_bar`)

.panelset[
.panel[.panel-name[The Code]


```{r fig.align='center', message=FALSE, warning=FALSE, out.width="50%"}
to_na <- function(x){ifelse(x<0,NA,x)}



anes16a <- anes16 %>%
  select(V161003, V162034) %>%
  mutate(across(everything(), to_na)) %>%
  mutate(V162034 =  as.factor(V162034),
         V161003a = case_when(
    V161003 == 1 ~ "Always",
    V161003 == 2 ~ "Most of the time",
    V161003 == 3 ~ "About half the time",
    V161003 == 4 ~ "Some of the time",
    V161003 == 5 ~ "Never"
  ))

# plot  
bar <- ggplot(anes16a, aes(x = V161003a)) + geom_bar()
```


]

.panel[.panel-name[Output]

```{r echo=FALSE, fig.align="center", fig.align='center', message=FALSE, warning=FALSE, out.width="50%"}
bar
```
]
]
  

---

# Histogram (`geom_histogram`)
.panelset[.panel[.panel-name[Code]
```{r fig.align='center', message=FALSE, warning=FALSE, out.width="50%"}

anes16b <- anes16 %>%
  mutate(V161267 = ifelse(V161267<0,NA,V161267))

hist <- ggplot(anes16b, aes(x = V161267)) + geom_histogram()

```
]
.panel[.panel-name[Output]
```{r, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, out.width="50%", echo=FALSE}
hist
```
]]
---

# Dot chart (`geom_dot`)
.panelset[.panel[.panel-name[Code]
```{r fig.align='center', message=FALSE, warning=FALSE, eval=TRUE}

# Age, FT of Trump, 3pt Party ID Interest
anes16c <- anes16 %>%
  select(V161092, V161267, V161155, V161003) %>%
  mutate(across(everything(), to_na))

dot <- ggplot(data = anes16c, aes(x = V161267, y = V161092 )) + geom_point()

```
]
.panel[.panel-name[Output]
```{r, fig.align='center', message=FALSE, warning=FALSE, out.width="50%", echo=FALSE}
dot
```
]
]

---

# Line Chart (`geom_smooth`)
.panelset[.panel[.panel-name[Code]
```{r fig.align='center', message=FALSE, warning=FALSE, out.width="50%"}
anes16ca <- anes16 %>%
  select(V161092, V161267) %>%
  mutate(across(everything(), to_na))

smooth <- ggplot(data = anes16ca, aes(x = V161267, y = V161092 )) + geom_smooth()

```
]
.panel[.panel-name[Output]
```{r,fig.align='center', message=FALSE, warning=FALSE,  out.width="50%", echo=FALSE}
smooth
```

]]
---

# Bonus! Point + Smooth

```{r, fig.align='center', message=FALSE, warning=FALSE, out.width="50%"}
dot + geom_smooth()

# ggplot(data = anes16c, aes(x = V161267, y = V161092 )) + geom_dot + geom_smooth()

```

---

# Line Chart (`geom_line`)
.panelset[.panel[.panel-name[Code]
```{r, }
anes16d <- anes16 %>%
  select(V161267, V161086, V162079) %>%
  rename(pre = V161086, post = V162079) %>%
  mutate(across(everything(),to_na)) %>%
  summarise(`Pre-Election` = mean(pre, na.rm = T), 
            `Post-Election` = mean(post, na.rm = T)) %>%
  pivot_longer(names_to = "xvar", cols = everything())

line <- ggplot(anes16d, aes(x = xvar, y = value)) + geom_line(group = "identity")
```
]
.panel[.panel-name[Output]
```{r, fig.align="center", out.width="50%"}
line
```


]]


---
# Color aesthetic (`color`)
.panelset[.panel[.panel-name[Code]
```{r}
anes16c <- anes16c %>%
  filter(!(V161155 == 5), !(V161155 == 0)) %>%
  mutate(V161155 = as.factor(V161155))

dot1 <- ggplot(data = anes16c, aes(x = V161267, y = V161092, color = V161155)) + geom_point()
```

]

.panel[.panel-name[Output]

```{r, fig.align="center", out.width="50%"}
dot1
```

]

]


---
# Dot size
.panelset[.panel[.panel-name[Code]
```{r}
anes16c <- anes16c %>%
  filter(!(V161155 == 5), !(V161155 == 0)) %>%
  mutate(V161155 = as.factor(V161155))

dot2 <- ggplot(data = anes16c, aes(x = V161267, y = V161092, size = V161003)) + geom_point()

```

]
.panel[.panel-name[Output]

```{r, fig.align="center", out.width="50%"}
dot2
```

]

]
 
---

# Transparency
.panelset[.panel[.panel-name[Code]
```{r}
anes16c <- anes16c %>%
  filter(!(V161155 == 5), !(V161155 == 0)) %>%
  mutate(V161155 = as.factor(V161155))

dot3 <- ggplot(data = anes16c, aes(x = V161267, y = V161092, alpha = V161003)) + geom_point()

```

]
.panel[.panel-name[Output]

```{r, fig.align="center", out.width="50%"}
dot3
```

]

]

---

# Grouping 
.panelset[.panel[.panel-name[Code]
```{r}
anes16a <- anes16 %>%
  select(V161003, V162034) %>%
  mutate(across(everything(), to_na)) %>%
  mutate(V162034 =  as.factor(V162034),
         V161003a = case_when(
    V161003 == 1 ~ "Always",
    V161003 == 2 ~ "Most of the time",
    V161003 == 3 ~ "About half the time",
    V161003 == 4 ~ "Some of the time",
    V161003 == 5 ~ "Never"
  ))

# plot  
bar1 <- ggplot(anes16a, aes(x = V161003a, fill=V162034)) + geom_bar()
```


]

.panel[.panel-name[Output]

```{r echo=FALSE, fig.align="center", fig.align='center', message=FALSE, warning=FALSE, out.width="50%"}
bar1
```
]]


---

# Helpful theme stuff:

- `theme_`
- `theme`
- `coord_flip`
- `scale_MAP_TYPE`
- `labs`