---
title: "Welcome to Election Data Science!"
subtitle: "Election Data Science"  
author: 
  - "Peter Licari"
date: '9/1/2020'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF",
  text_font_google = google_font("Poppins")
)
style_duo_accent
xaringanExtra::use_xaringan_extra()
```

# Welcome to the Course!

--
- Learn a little about me.

--
- Learn a little about you
--

- Learn a little about the course.


---
# Slide with all the important stuff on it

- My name: Peter Licari
  - Call me "Mr", "Professor", "Peter"-whatever (just not **Dr** - except on October 13th)
- Class times: Tuesday (*8:30-10:25*); Thursday (*9:35-10:25*)
- Email: [plicari13@ufl.edu](mailto:plicari13@ufl.edu)
- Office hours: Wednesday (*8:00-10:00* or by appointment)
---

# A little about me
--

- 6th year PhD candidate in Political Science
  - Dissertation looks at how video games increase political participation
--

- Got my B.A. in Political Science at UNF, M.A. here at UF.
--

- Data Scientist at Morning Consult (As of, like, literally yesterday `r emo::ji("geek")`)
--

- Dog dad, cat dad, and *dad* dad.
  - Dude, Asia, and Rosalina, respectively.
--

- My teams...
---
# A little about you

- Name
- Major
- What made you interested in this course
- What lady luck decides...
  1. Book
  2. Movie
  3. Sports Team
  4. TV Show (YouTube Channel is acceptable)
  5. Game
  6. Non-U.S. Media 

???
1. Once a Runner
2. Without Limits
3. San Fran
4. Avatar the Last Air Bender/ Evan and Kaitlyn
5. Scrabble/Fallout New Vegas.
6. My Hero Academia.
---
# Assignments
Class is scored out of 1,000 points.
- **Attendance:** 50 points (5%)
- **Check-In Quizzes:** 50 points (5%)
- **Weekly Labs:** 300 points (30%)
- **Blog Posts:** 200 points (20%)
- **Final Project:** 400 points (40%)
  - *Proposal:* 50 points
  - *Check-in:* 50 points
  - *Technical Report:* 150 points
  - *Public-Facing Output:* 150 points
  
???
The syllabus details some extra credit opportunities.
Give R bloggers as an example. Use your doggo one for the project.
---
# Course Policies

- Office Hours
- Attendance
- Late assignments
- Children/Disruptions in class
- Demeanor
- Technology

---
# What to expect in a given week.
- **Tuesday**
  - Lab
  - Conceptual Lecture
  - Practical Lecture/Tutorial
- **Thursday**
  - Practical lab.
  - Work in break-out groups

- Also, lots of memes.